The Unintended Consequences of Energy Reduction

### 2024-02-15 paper/paper.qmd

As a response to the growing public outcry, the government of Japan took swift action to significantly reduce nuclear power generation by implementing a new energy policy to eliminate nuclear power by 2030 [@intechopen]. Before the accident, nuclear energy comprised 25% of Japan’s energy market [@intechopen]. After the accident, the market share stayed under 1.7% until 2016 [@intechopen]. The significant and abrupt reduction in overall power generation caused Japan to steer towards potential countrywide power shortages [@guardian]. To combat this, the government pushed a power saving campaign to the public [@guardian]. In fear of blackouts, energy saving was quickly adopted by Japanese citizens, resulting in a considerable decrease in electricity consumption [@intechopen].

The energy saving campaign was not the sole factor in the sharp decline in energy consumption as the price for electricity also surged. As many nuclear reactors were shut down, fossil fuels grew to dominate Japan’s energy market. In 2011, oil usage rose by an estimated 85% while natural gas increased by 25% [@energyPolicyYearLater]. This shift to predominate fossil fuel usage had a substantial impact on the price and consumption of electricity as the majority of fossil fuels utilized in Japan are imported from other countries, resulting in expensive prices [@energyPolicyYearLater]. Relative to the mean price of electricity in 2011, the price of electricity had increased by over 10% in 2013 and further increased by nearly 20% in 2014 [@energyPolicyYearLater]. Consequently, electricity consumption decreased each year as the commodity became less affordable. The Japanese government was successful in reducing energy consumption; however, their energy-saving policies and campaigns cause potential detrimental consequences to the public.

In this paper, we follow He and Tanaka’s paper to produce a reproduction of their findings and discuss further about Japan’s future as their government reverses their stance on nuclear energy and begins restarting the use of old power plants as a response to fossil fuel shortages and the public pressure to reduce the country’s carbon footprint. We analyze the major reduction in energy consumption from 2005 to 2015 caused by the decline in nuclear energy usage, the sharp increase in electricity prices and the energy saving campaign pushed by the Japanese government to investigate their impact on mortality rate in high temperatures. We find that the mortality rate increases as energy consumption decreases. An important finding is that air conditioner usage plays a vital role in preventing deaths related to high temperatures. Therefore, the pressure to conserve energy and hence not utilize air conditioners, increases the mortality rate in Japan.

The remainder of this paper is structured as follows: the Data section details the collection
and processing of nuclear energy production, energy consumption and energy price, and temperature-mortality data. Additionally, the section includes discussion on the measurement of the original dataset. The following Results section displays and analyzes tables
and figures that illustrate the trends found in the data. The Discussion section draws conclusions from those tables and figures and details the importance of our findings and weaknesses of our paper.

@@

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

>>>>

As a response to the growing public outcry, the government of Japan took swift action to significantly reduce nuclear power generation by implementing a new energy policy to eliminate nuclear power by 2030 [@intechopen]. Before the accident, nuclear energy comprised 25% of Japan’s energy market [@intechopen]. After the accident, its market share stayed under 1.7% until 2016 [@intechopen]. The significant and abrupt reduction in overall power generation caused Japan to steer dangerously close to potential countrywide power shortages [@guardian]. To combat this, the government pushed a power saving campaign to the public [@guardian]. In fear of blackouts, energy saving was quickly adopted by Japanese citizens, resulting in a considerable decrease in electricity consumption [@intechopen].

We find that the energy saving campaign was not the sole factor in the sharp decline in energy consumption as the price of electricity also surged. As many nuclear reactors were shut down, fossil fuels grew to dominate Japan’s energy market. In 2011, oil usage rose by an estimated 85% while natural gas increased by 25% [@energyPolicyYearLater]. This shift to predominately fossil fuel usage had a substantial impact on the price and consumption of electricity as the majority of fossil fuels utilized in Japan are imported from other countries, resulting in expensive prices [@energyPolicyYearLater]. Relative to the mean price of electricity in 2011, the price of electricity increased by over 10% in 2013 and further increased by nearly 20% in 2014 [@energyPolicyYearLater]. Consequently, electricity consumption decreased each year as the commodity became less affordable. The Japanese government was successful in reducing energy consumption; however, their energy-saving policies and campaigns may cause potential detrimental consequences to the public.

In this paper, we follow He and Tanaka’s paper (2019) to produce a reproduction of their findings and discuss further about Japan’s future as their government reverses their stance on nuclear energy and begins restarting the use of old power plants as a response to fossil fuel shortages and the public pressure to reduce the country’s carbon footprint. We analyze the major reduction in energy consumption from 2005 to 2015 caused by the decline in nuclear energy usage, the sharp increase in electricity prices, and the energy saving campaign pushed by the Japanese government in order to investigate their impact on the mortality rate related to high temperatures. We find that the mortality rate increases as energy consumption decreases. An important finding is that air conditioner usage plays a vital role in preventing deaths related to high temperatures. Therefore, the pressure to conserve energy and hence not utilize air conditioners, increases the mortality rate in Japan.

The remainder of this paper is structured as follows: the Data section details the collection
and processing of nuclear energy production, energy consumption, energy price, and temperature-mortality data. Additionally, the section includes discussion on the measurement of the original dataset. The following Results section displays and analyzes tables
and figures that illustrate the trends found in the data. The Discussion section draws conclusions from those tables and figures and details the importance of our findings and weaknesses of our paper.

@@

Our findings give insight into the consequences of energy saving in Japan, particularly the potential fatalities that may occur. It is also important to discuss the growing use of fossil fuel dependent energy generation as a response to the Fukushima nuclear accident and its devastation on the environment.

## Findings

After the Fukushima accident, the Japanese public’s trust of nuclear power was retracted and the use of nuclear power began to diminish. Figure 1 emphasizes this sharp decline in nuclear power energy production in Japan. As a consequence of the decline in nuclear energy production, Figure 2 shows the extreme increase in energy prices and the decline in energy consumption. These findings can be attributed to the Japanese government urging the public to use less energy and the expensive cost of fossil fuel use for energy production. Figure 3 demonstrates that the energy saving campaign orchestrated by the government and the use of expensive fossil fuels can have harmful effects. Specifically, Figure 3 illustrates the potential increase in the mortality rate of Japanese citizens due to their reduction in energy consumption. In particular, Barreca et al. (2016) states that air condition usage reduces the mortality rate related to high temperatures by a significant amount [@barreca]. By utilizing their findings and He and Tanaka’s (2019) findings, we draw the conclusion that the mortality rate increases by 0.8% during temperatures between 25°C and 30°C and by 0.19% during temperatures over 30°C [@heTanaka]. This is significant as it demonstrates that the push for energy saving may cause fatalities in high temperature environments. This consequence of energy saving can be further illustrated by Figure 4 and Figure 5, whereby the number of heatstroke incidents sharply increased from 2010 to 2011.

## Japan’s Relationship with Nuclear Energy and its Impact on the Public

As shown in Figure 1, the use of nuclear energy is slowly becoming obsolete in Japan in light of the Fukushima accident. However, in 2022, the Japanese government announced a plan to revitalize the use of nuclear energy [@npr]. This decision comes as a potential solution to fossil fuel shortages and the push for low carbon emissions. Nuclear energy production produces little waste, is less expensive to operate, and generates low carbon emissions [@energy]. The government plans on maximizing the output of all existing nuclear power plants and they aim to build new reactors, proclaiming that nuclear energy will contribute to 20-22% of the country’s energy output by 2030 [@npr]. The government reversed the nuclear energy policy, but low public demand for nuclear energy persists. Thus, energy prices will remain high as individuals continue to advocate against nuclear energy usage and turn to fossil fuels. Additionally, there are barriers to reinstating nuclear power generation that exacerbate the trend found in Figure 3. One barrier is the Nuclear Regulation Authority which enforced stricter safety regulations for reactors [@intechopen]. Restarts of existing nuclear reactors have been slow due to these stricter regulations put in place following the Fukushima accident [@npr]. Additionally, the government’s new stance on nuclear energy does not come without opposition by the public. One survivor of the Fukushima disaster voices their frustration, “The Fukushima disaster is not over yet and the government seems to have already forgotten what happened.” [@npr]. Furthermore, citizens can bring forth injunctions to prevent the operation of nuclear reactors. After the Fukushima tragedy, many courts in Japan may grant these injunctions to protect the public's safety [@intechopen]. Thus, as the return to nuclear energy is met with resistance, the use of fossil fuels persists; thereby causing the price of energy to increase and the consumption of energy to decrease, leading to potentially more fatalities in high temperatures.

## Future Research Proceedings

For future research, more data should be collected regarding deaths due to high temperatures to investigate its relationship with energy prices. Additionally, as global temperatures continue to soar each year, it is imperative that more environmentally friendly solutions to energy generation be adopted.
